### 2018-04-11 paper.Rmd
color example

To demonstrate how one might discover such a non-stationary, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer [see @wagner2013]. However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary -- e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways -- first using the childesb shiny apps, and then using the `childesr` package.

@@

To investigate whether the frequency of color words is stationary over development, a user can navigate to the Frequency app, and enter a set of color words into the `Word` selector separated by a comma: here "blue, red, green." Because the question of interest is about the frequency of words in the input (rather than produced by children), the `Speaker` field can be set to reflect this. In this example we select "Mother" because, across corpora in CHILDES, the speaker tagged with the Mother role tends to produce most of the words that children hear. Because children learn most of their basic color words by the age of 5, the age range of 1-5 years is a reaonable choice for `Ages to include`. The results of these selections are shown in Figure \ref{fig:color-shiny}. We can also create a hyperlink to store these set of choices so that we can share these results with others, or with ourselves in the future, by clicking on the `Bookmark` button in the bottom left corner.

@@

```{r color-shiny, fig.width = 10, fig.height = 6, fig.cap = "An example of using the Frequency shiny app to explore how children's color input changes over development"}
knitr::include_graphics("images/color-shiny.png")
```

@@

Figure \ref{fig:color-plot}.

```{r color-plot, fig.height = 5, fig.width = 10, fig.cap ="Color frequency as a function of age. Points represent means across transcripts, error bars represent 95% confidence intervals computed by nonparametric bootstrap", cache = TRUE, message = FALSE, warning = FALSE}
colors <- c("black", "white", "red", "green", "yellow", "blue", "brown", 
             "orange", "pink", "purple")

graph_colors <- data_frame(color = colors,
                           graph_color = if_else(color == "white", 
                                                 "light gray", color))

color_counts <- get_types(collection = "Eng-NA", role = c("Mother", "Father"),
                          age = c(12,60), type = colors)

parents <- color_counts %>%
  distinct(collection_id, corpus_id, transcript_id, speaker_id) %>%
  left_join(get_speaker_statistics(collection = "Eng-NA")) %>%
  select(collection_id, corpus_id, transcript_id, speaker_id, num_tokens)

count_estimates <- color_counts %>%
  left_join(parents) %>%
  mutate(age = target_child_age / 30.5,
         age_bin = as.integer(floor(age / 6) * 6),
         color = tolower(gloss)) %>%
  group_by(age_bin, color, transcript_id) %>%
  summarise(count = sum(count), tokens = sum(num_tokens)) %>%
  mutate(prop = count/tokens * 1000000) %>%
  tidyboot::tidyboot_mean(prop) %>%
  left_join(graph_colors) %>%
  mutate(color = factor(color, levels = colors))

ggplot(count_estimates, aes(x = age_bin, y = empirical_stat, 
                            ymin = ci_lower, ymax = ci_upper,
                            color = color, label = color)) + 
  geom_pointrange(position = position_dodge(.5)) + 
  geom_point() +
  geom_smooth(method = "loess", se = F, span = .5) +
  facet_wrap(~ color, scales = "free_y") + 
  theme_classic(base_size = 14) +
  scale_color_manual(values = graph_colors$graph_color) +
  scale_x_continuous(breaks = seq(12, 60, 12),
                     name = "Age (months)") + 
  scale_y_continuous(name = "Frequency (per million words)") +
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) + 
  theme(legend.position = "none")
```

non-stationarity

>>>>

To demonstrate how one might discover such a non-stationary, we take as a case study the frequency with which children hear the color words of English (e.g. "blue", "green"). Color words tend to be learned relatively late by children, potentially in part due to the abstractness of the meanings to which they refer [see @wagner2013]. However, within the set of color words, the frequency with which these words are heard predicts a significant fraction of the variance in their order of acquisition [@yurovsky2015]. But are these frequencies stationary -- e.g. do children hear "blue" as often at 12 months as they do at 24 months? We answer this question in two ways -- first using the web apps, and then using the `childesr` package.

@@

To investigate whether the frequency of color words is stationary over development, a user can navigate to the Frequency app, and enter a set of color words into the `Word` selector separated by a comma: here "blue, red, green." Because the question of interest is about the frequency of words in the input (rather than produced by children), the `Speaker` field can be set to reflect this. In this example we select "Mother" because, across corpora in CHILDES, the speaker tagged with the Mother role tends to produce most of the words that children hear. Because children learn most of their basic color words by the age of 5, the age range of 1-5 years is a reasonable choice for `Ages to include`. The results of these selections are shown in Figure \ref{fig:color-shiny}. We can also create a hyperlink to store these set of choices so that we can share these results with others, or with ourselves in the future, by clicking on the `Bookmark` button in the bottom left corner.

@@

```{r color-shiny, out.width = "500px", fig.cap = "An example of using the Frequency shiny app to explore how children's color input changes over development"}
knitr::include_graphics("images/color-shiny.png")
```

@@

We can analyze these learning trajectories using childesr by braking the process into five steps: (1) define our words of interest, (2) find the frequencies with with which children hear these words, (3) Find the proportion of the *total words* children hear that these frequencies account for, (4) Aggregate across transcripts and children to determine the error in our estimates of these proportions, and (5) plot the results.

For this analysis, we will define our words of interest as the basic color words of English (except for gray, which children hear very rarely). We store these in the `colors` variable, and then use the `get_types` function from `childesr` to get the type frequency of each of these words in all of the corpora in CHILDES. For demonstration, we look only at the types produced by the speakers in each corpus oftagged as Mother and Father. We also restrict ourselves to children from 1-5 years old (12 to 60 months), and look only the American English corpora.

```{r define-colors, echo = TRUE}
colors <- c("black", "white", "red", "green", "yellow", "blue", "brown", 
             "orange", "pink", "purple")

color_counts <- get_types(collection = "Eng-NA", role = c("Mother", "Father"),
                          age = c(12,60), type = colors)
```

```{r graph-colors}
graph_colors <- data_frame(color = colors,
                           graph_color = if_else(color == "white", 
                                                 "light gray", color))

```

To normalize correctly (i.e. to ask what proprotion of the input children hear consists of these color words), we need to know how many total words these children hear from their parents in these transcripts. To do this, we use the `get_speaker_statistics` function, which will return a total number of tokens (`num_tokens`) for each of these speakers.

```{r get-token-frequencies, echo = TRUE}
# Get the ids corresponding to all of the speakers we are interested in
parent_ids <- color_counts %>%
  distinct(collection_id, corpus_id, transcript_id, speaker_id)

# Find in the total number of tokens produced by these speakers
parents <- parent_ids %>%
  left_join(get_speaker_statistics(collection = "Eng-NA")) %>%
  select(collection_id, corpus_id, transcript_id, speaker_id, num_tokens)
```

We now join these two pieces of information together--how many times each speaker produced each color word, and how many total words they produced. We then group the data into 6-month age bins, and compute the proportion of tokens that comprise each color for each child in each 6-month bin. For comparability with the web app analysis, we convert the proportions to parts per million words
```{r join-colors-and-tokens, echo = TRUE}
count_estimates <- color_counts %>%
  left_join(parents) %>%
  mutate(age_months = target_child_age / 30.5,
         age_bin = as.integer(floor(age_months / 6) * 6),
         color = tolower(gloss)) %>%
  group_by(age_bin, color, target_child_id, transcript_id) %>%
  summarise(count = sum(count), num_tokens = sum(num_tokens)) %>%
  summarise(count = sum(count), num_tokens = sum(num_tokens)) %>%
  mutate(parts = count / num_tokens * 1e6)
```

Finally, we use non-parametric bootstrapping to estimate 95% confidence intervals for our estimates of the parts per million words of each color term with the `tidyboot` package.

```{r computer-color-error, echo = TRUE}
count_estimates_with_error <- count_estimates %>%  
  tidyboot::tidyboot_mean(parts) %>%
  left_join(graph_colors) %>%
  mutate(color = factor(color, levels = colors))
```

Figure \ref{fig:color-plot} shows the results of these analyses: Input frequency varies significantly over the 1-5 year range for nearly every color word.

```{r color-plot, fig.height = 5, fig.width = 10, fig.cap ="Color frequency as a function of age. Points represent means across transcripts, error bars represent 95% confidence intervals computed by nonparametric bootstrap", cache = TRUE, message = FALSE, warning = FALSE}
ggplot(count_estimates_with_error, aes(x = age_bin, y = empirical_stat, 
                            ymin = ci_lower, ymax = ci_upper,
                            color = color, label = color)) + 
  geom_pointrange(position = position_dodge(.5)) + 
  geom_point() +
  geom_smooth(method = "loess", se = F, span = .5) +
  facet_wrap(~ color, scales = "free_y") + 
  theme_classic(base_size = 14) +
  scale_color_manual(values = graph_colors$graph_color) +
  scale_x_continuous(breaks = seq(12, 60, 12),
                     name = "Age (months)") + 
  scale_y_continuous(name = "Frequency (per million words)") +
  geom_dl(method = list(dl.trans(x=x +.2), "last.qp", cex=1)) + 
  theme(legend.position = "none")
```
